### 2017-08-09 doc/MSc-dissertation.tex
DNS

\section{MHD turbulence}

@@

Afterwards, equations for the equal time two-point correlation functions of the velocity and magnetic fields allowed him to derive an equation for the two-point correlation function of the magnetic field for non-helical HIT. Using that, a Schroedinger-type time independent equation arose. These showed the possibility of bound states, provided a critical magnetic Reynolds number $Re_M^{crit}$ within the range [30:60] was reached, which meant that a small-scale dynamo was being excited~\cite{brandenburg2005astrophysical}.

@@

\chapter{Design and/or development (of my project)}

@@

\chapter{Stuff that's too detailed}

Appendices should contain all the material which is considered too
detailed to be included in the main body of the text, but which is
important enough to be included in the thesis.

Perhaps this is a good place to mention \BibTeX.

You can do references in the simple way explained in the introduction,
or you can use \BibTeX.

\section{\BibTeX}
\label{sec:bibtex}

It is convenient to use \BibTeX\ to compile your bibliography.  First
you need to create a .bib file e.g.  you may call it ref.bib Then you
can put all your references into the file with entries such as
\begin{verbatim}
@Book{ob:bornwolf,
     author = "Born, M and Wolf, E",
     title  = "Principles of Optics",
     publisher = "Cambridge University Press",
     year = 1999,
     edition = {7th},
}

@Article{jr:ashkin,
Author = {A. Ashkin and J.M. Dziedzic and J.E. Bjorkholm and S. Chu},
Title = "Observation of a single beam gradient force optical tap for 
dielectric particles",
Journal = "Optics Letters",
Volume = 11,
Pages = "288-290",
Year = 1986}

@INPROCEEDINGS{seger,
 author = {J. Seger and H.J. Brockman},
 title = {What is bet-hedging?},
 editors={P.H. Harvey and L. Partridge},
 booktitle = {Oxford Surveys in Evolutionary Biology},
 year={1987},
 page={18},
 publisher={Oxford University Press},
 place={Oxford}}
\end{verbatim}
for a book, an article in a journal or an article in a proceedings volume
respectively.

Inside your \LaTeX\ file
you should include 
\begin{verbatim}
\bibliographystyle{unsrt}                      
and
\bibliography{ref}
\end{verbatim}
The first command determines the reference style, here plain and 
unsorted. With this referencing style 
a numerical referencing system (which is now the most
common in physics literature) is used and the numbering of references
will be the order in which they appear in the document. Alternatively, 
you could use
a customised `style file' but there is no real need.  The second
command just inputs your .bib file Note that only the references cited
in the text will appear in the bibliography so you can have spare
references in your .bib file.

You use the name you have given to an entry (e.g.
for the book example above the name is ob:bornwolf)
to cite the relevant article
by using the cite command in your \LaTeX\ file e.g. 
\begin{verbatim}
\cite{ob:bornwolf}
\end{verbatim}

\section{Producing your documents using \texttt{pdflatex}}

To use pdflatex your figures need to be in pdf format.  You can convert almost any image file to pdf using \texttt{convert}.  e.g. \texttt{convert myfigure.png myfigure.pdf}.

The first time you should type:
\begin{verbatim}
  pdflatex ProjectReport
  bibtex ProjectReport
  pdflatex ProjectReport
  pdflatex ProjectReport
\end{verbatim} 
This first time you run\texttt{pdflatex} it will produce a
\texttt{ProjectReport.aux}.  The \BibTeX\ command reads in the
bibliography file and makes the files \texttt{ProjectReport.bbl} and
\texttt{ProjectReport.blg} files.  These files are read in the next
\texttt{pdflatex} command, but you'll still have ``undefined
cross-reference'' errors which are sorted out by the last
\texttt{pdflatex} command.

Subsequently, you should only need to do one (or two)
\texttt{pdflatex}s, or \texttt{pdfbibtex} followed by
\texttt{pdflatex} twice if you change any references.

\vspace{5mm} You may also use plain \texttt{latex} instead of
\texttt{pdflatex}.  This requires you to use postscript graphics
instead of pdf.

>>>>

\section{MHD turbulence}
\label{sec3.2}

@@

Afterwards, equations for the equal time two-point correlation functions of the velocity and magnetic fields allowed him to derive an equation for the two-point correlation function of the magnetic field for non-helical HIT. Using that, a Schroedinger-type time independent equation arose. These showed the possibility of bound states, provided a critical magnetic Reynolds number $Re_M^{crit}$ within the range [30:60] was reached, which meant that a small-scale dynamo was being excited~\cite{brandenburg2005astrophysical, kazantsev1968enhancement}. In Kazantsev's theory, the aforementioned critical magnetic Reynolds number needs to arise in a situation when $\nu \geq \eta$, i.e. when the magnetic Prandtl number greater or equal than unity~\cite{kazantsev1968enhancement}. The value of $Re_M^{crit}$ can be decreased by injecting kinetic helicity into the problem--this also makes the magnetic energy to be placed mainly in the large scales, i.e. achieves large-scale dynamo action, as a consequence of the $\alpha$-effect~\cite{leorat1981fully}.

Recently, numerical simulations of forced, non-helical, incompressible HIT in MHD have suggested that there is a dependence on the kinetic Reynolds number for the critical magnetic Reynolds number. Particularly, that dynamo action can be ``shut down if the magnetic Prandtl number $Pr_M = Re_M/Re$ is less than some critical value $Pr_M^{crit.} \lesssim 1$ even for $Re_M$ for which dynamo exists at $Pr_M \geq 1$''~\cite{schekochihin2004critical}. This can be observed by keeping the magnetic diffusivity constant and varying the magnetic Prandtl number for an $\eta$ that maintains dynamo action when $Pr_M\gtrsim1$.

@@

\chapter{Direct numerical simulations}

Having stated \textit{some} important theoretical results regarding homogeneous isotropic turbulence (HIT) for magnetohydrodynamics (MHD), like the existence of a critical magnetic Reynolds number for small-scale dynamo action, it is important to note that the main technique for investigating turbulence is through numerical simulations. This has to do with the fact that analytic solutions for the MHD equations~(\ref{eq3.1}-~\ref{eq3.3}), even for the simplest turbulent flows, do not exist~\cite{moin1998direct}. Different numerical methods have been developed to analyse MHD turbulence depending on the specific problem that is being studied. For example, if the smaller length scales relevant to the problem can be neglected, one might want to use Reynolds-averaged Navier–Stokes (RANS) or Large Eddy Simulations (LES) for a slightly more complex description. The main motivation for choosing RANS or LES is that resolving the smaller length scales is very computationally expensive and not necessary for certain problems. However, if all the length scales in the problem need to be resolved, then Direct Numerical Simulations (DNS) ought to be used.

As far as numerical methods for solving (magneto)hydrodynamic equations go, DNS is by far the most complex and computationally expensive. As a consequence, DNS was very limited in the size of simulations that it could preform in its early stages\footnote{The size of a DNS simulation determines the wave numbers (length scales) that it can achieve, and thus the Reynolds number that can be simulated. The number of operations required for DNS scales with $Re^{9/4}$~\cite{mccomb1990physics}.}. Direct numerical simulations were developed by Orszag and Patterson in the 1970s, where they simulated 3D hydrodynamic isotropic turbulence in a periodic box of $32^3$ grid points with periodic boundary conditions~\cite{orszag1970analytical}. Around $50$ years after their pioneering work, and with the use of paralellisation and graphic card units (GPUs), the largest DNS simulation is one of $4096^3$ grid points~\cite{yokota2013petascale}. A brief account of what is DNS and how codes that use it are set up is initially provided.

\section{DNS basics}

Broadly, DNS is a type of simulation that solves the Navier-Stokes equations for hydrodynamics, and the Maxwell equations as well for MHD, without the use of turbulence models. Alternatively, DNS resolves all the spatial and temporal scales related to a specific problem. In order to solve the Navier-Stokes equations, DNS uses a pseudo-spectral method. A brief description of a DNS procedure is provided in this section.

The main problem with computational fluid dynamics is the nonlinear term $(\bm u \cdot \bm \nabla) \bm u$ in the Navier-Stokes (and MHD) equations. As it was mentioned previously, this term is directly related to turbulence in a system. In order to reformulate the Navier-Stokes equations to have only one dynamic variable, the velocity field, it is necessary to Fourier transform (FT) them and to use the incompressibility condition $\dot{p}=\bm \nabla \cdot \bm u = 0$ (or an equation of state $p=p(\rho)$)\footnote{For MHD two dynamic variables will exist, the velocity and the magnetic fields.}. This means to set the pressure term as a function of the velocity field. However, the nonlinear term remains in the new equations and now it is a convolution over all possible wave numbers~\cite{mccomb1990physics}!

After generating a velocity field, the Fourier-transformed Navier-Stokes equations are then numerically solved. It is worth noting that the nonlinear convolution alone will require $\mathcal{O}(N^6)$ operations~\cite{mccomb1990physics}. This will scale horribly when dealing with large boxes. But wait, that's $\mathcal{O}(N^6)$ operations for the nonlinear term per time step! In order to mitigate this problem, the convolution theorem is used. The convolution theorem states that a convolution's Fourier transform $\mathcal{F}\{f \ast g\}$ is defined by the point-wise product of the Fourier transform of its components $\mathcal{F}\{f\} \cdot \mathcal{F}\{g\}$. This means that the convolution can be described as $f \ast g = \mathcal{F}^{-1}\{ \mathcal{F}\{f\} \cdot \mathcal{F}\{g\} \}$. Using fast Fourier transforms (FFT), and considering that for the nonlinear term in the Navier-Stokes equation both $f$ and $g$ are the velocity field $\bm u (\bm k, t)$, the computational complexity for obtaining the nonlinear term is decreased to $\mathcal{O}(N^3 \log_2 N)$~\cite{mccomb1990physics}.

The pseudo-spectral method does help in decreasing the number of operations required to obtain the nonlinear term. However, this needs to be done in every single time step (and for two fields in MHD). Exactly what time stepping strategy is used, that depends on different codes. As far as the actual value of the time step, each code will have a Courant–Friedrichs–Lewy (CFL) condition to make sure the results do not diverge. For the MHD code that was used for the simulations shown later, a forward in time Heun's method was used and the CFL condition was $dt < 0.1/(u_{rms}k_{max})$. This means that $dt$ is inversely proportional to the box size, which means that larger boxes will need smaller time steps. It is clear how expensive it is to get simulations that resolve all length scales, and why making simulations in larger boxes is very time consuming~\cite{yoffe2013investigation, LinkmannMoritzFrederikLeon2016Spim}.

What was mentioned above is simply a short introduction to DNS, and it is considerably more complicated than what was shown. However, for the purpose of this work no more details will be given on the additional intricacies of DNS. Given that in the present work the onset of small-scale dynamo action is being investigated it is of great importance to resolve the smallest length scales of the problem. This means that the type of numerical simulation required is DNS. Keeping in mind that running large boxes with DNS is computationally expensive (and thus financially expensive) it is worth mentioning how to choose the appropriate box sizes for a given set of initial parameters.

\section{DNS resolution}

As it was mentioned in sections~\ref{sec2.4}~and~\ref{sec3.2}, problems with turbulence have various length scales associated with different physical processes. For simulations with forcing the largest length scale is commonly the one at which energy is input into the system.

Although Mouschovias had already published
by the mid 1970s 2D static models of magnetically supported clouds, it was not
until 1980 that the first dynamic MHD simulation was published~\cite{norman2010historical}.

Results!

A critical magnetic Reynolds number was found to be $Re_M^{crit}=55\pm5$ for simulations with $Pr_M\gtrsim1$, which is very close to a reported value of $Re_M^{crit}\sim60$ (also for $Pr_M\gtrsim1$) in~\cite{schekochihin2007fluctuation}.

\chapter{Critical magnetic Prandtl number for small-scale dynamos}

@@
